---
output:
  html_document: default
  word_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

title: "PKpara and Effect estimation" author: "Vivek Sahebrao LOhar" date: "2022-11-07" output: html_document: df_print: paged ---

# Introduction:

# Data:

```{r,,echo=FALSE}
rm(list = ls(all=TRUE))
data=read.csv("E:/PRACTICAL SEM III/data set/BABEData.csv",header = F)
```

#Extract:

```{r,echo=FALSE}
ti=data[1,5:20]
ci=data[2:49,5:20]
```

# defining the pkpara function:

-   $t_{max}$ : maximum time point.
-   $C_{max}$ : maximum concentration.
-   $AUC_{t}$ : area under curve at time point t.
-   $AUC{oo}$ : area under curve at time point inf.
-   $t_{1/2}$ : elimination half time.
-   $k_{e}$ : rate constant.

```{r,echo=FALSE}
pkpara=function(ti,ci){
  pkpara=data.frame("tmax"=0,"cmax"=0,"auct"=0,"aucinf"=0,"thalf"=0,"ke"=0)
  n=length(ti)
  pkpara$cmax=max(ci)     # maximum concentration
  a=which.max(ci)
  pkpara$tmax=ti[a]       # time at which concentration is maximum
  pkpara$auct=sum((ci[2:n]+ci[1:(n-1)])*(ti[2:n]-ti[1:(n-1)]))  # partial auc truncated at the population median of t max 
  pkpara$ke=-2.303*coefficients(lm(log10(ci[(a+1):n])~ti[(a+1):n]))[2] #rate constant
  pkpara$thalf=0.693/pkpara$ke   #elimination half time 
  pkpara$aucinf=pkpara$auct+ci[n]/pkpara$ke
  return(pkpara)
}
```

# PKparameter estimation for one time point and one concentration for our dataset:

```{r,echo=FALSE}
out=pkpara(as.numeric(unlist(ti)),as.numeric(unlist(ci[1,]))) #Unlist tmax and cmax and store in 'out'.
out
```

#pkparameter estimation for whole data :

```{r,echo=FALSE}
for(i in 2:48){
  out[i,]=pkpara(as.numeric(unlist(ti)),as.numeric(unlist(ci[i,]))) #unlist tmax and cmax for whole dataset and store in out.
}
out
```

#To combine four columns of our data and their pkparameters:

```{r,echo=FALSE}
df=cbind(data[2:49,1:4],out)
print(df)
```

# "Five point summary" of pharmacokinetic parameter:

```{r,echo=FALSE}
summary(out)
```

#To calculate anova for the standard 2x2 crossover design.
# Statistical inferences for the fixed effects: - here we calculate carryover, drug formulation effect, period effect.
- $c_{hat}$: carryover effect (estimated).
- $f_{hat}$: linear contrast of the seq by period means.
- $p_{hat}$: unbiased estimator of period effect.

```{r,echo=FALSE}
#datasets:
data_1=read.csv("E:/PRACTICAL SEM III/data set/BABEData22.csv")
```

```{r,echo=FALSE}
n=length(data_1$Y)
y11=data_1[data_1$Per==1&data_1$Seq==1,1];n1=length(y11)
y12=data_1[data_1$Per==1&data_1$Seq==2,1]
y21=data_1[data_1$Per==2&data_1$Seq==1,1]
y22=data_1[data_1$Per==2&data_1$Seq==2,1];n2=length(y22)
ss_total=((n-1)/n)*var(data_1$Y)
df=data.frame(c(y11,y12),c(y21,y22))
ss_within=sum(apply(df,1,var))
n1=length(y11)
n2=length(y22)
ss_between=2*sum((apply(df,1,mean)-mean(data_1$Y))^2)
ss_total=(2*(n-1)/n)*var(data_1$Y)
chat=(mean(y12)+mean(y22))-(mean(y11)+mean(y21))
sscarry=2*n1*n2/(n1+n2)*chat^2/4
ssinter=ss_between-sscarry
fhat=(((mean(y21)-mean(y11))-(mean(y22)-mean(y12)))/2)^2
ssdrug=2*n1*n2/(n1+n2)*fhat

phat=(((mean(y21)-mean(y11))-(mean(y12)-mean(y22)))/2)^2
ssperiod=2*n1*n2/(n1+n2)*phat

ssintra=ss_within-ssdrug-ssperiod
source1=c("carry","inter","drug","period","intra","total")
df=c(1,(n1+n2-2),1,1,(n1+n2-2),2*(n1+n2)-1)
ss=c(sscarry,ssinter,ssdrug,ssperiod,ssintra,ss_total)
msq=ss/df
fcal=rep(0,6)
fcal[1]=msq[1]/msq[2]
fcal[2:4]=msq[2:4]/msq[5]
data.frame("source"=source1,"df"=df,"sumofsq"=ss,round(msq,3),round(fcal,4))

```

# Interpretation :

-   carryover effect : $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:C_T-C_R$ not equal to 0 here from p_value we fail to reject $H_{01}$ at 5% level of significance (i.e. Carryover effect of Test drug and Reference drug is same).

-   Drug formulation effect : $H_{02}:{F_T}-{F_R}=0$ vs $H_{12}:{F_T}-{F_R}\neq0$, here from p_value we fail to reject $H_{02}$ at 5% level of significance (i.e. Drug formulation effect of Test drug and Reference drug is same).

-   Period effect: $H_{03}:{P_1}-{P_2}=0$ vs $H_{13}:{P_1}-{P_2}\neq0$, here from p_value we fail to reject $H_{03}$ at 5% level of significance (i.e. Period effect of Test drug and Reference drug is same).

```{r}
#datasets:
data_1=read.csv("E:/PRACTICAL SEM III/data set/BABEData22.csv")
```

```{r,echo=FALSE}
crosanova=function(Y,Seq,Sub,Per){
  df=data.frame(Y,Seq,Sub,Per)
  Y=df[,1]
  n=length(Y)
  sst=((n-1)/n)*var(Y)
  y11=df[df$Per==1&df$Seq==1,1];n1=length(y11)
  y12=df[df$Per==1&df$Seq==2,1]
  y21=df[df$Per==2&df$Seq==1,1]
  y22=df[df$Per==2&df$Seq==2,1];n2=length(y22)
  df1=data.frame(y11,y21)
  df2=data.frame(y12,y22)
  v1=apply(df1,1,var)
  v2=apply(df2,1,var)
  sswit=sum(v1)+sum(v2)
  m1=sum((apply(df1,1,mean)-mean(Y))^2)
  m2=sum((apply(df2,1,mean)-mean(Y))^2)
  ssbet=2*(m1+m2)
  u1=y11+y21;u2=y12+y22
  d1=(y21-y11)/2;d2=(y22-y12)/2
  o1=d1;o2=d2
  
  chat=mean(u2)-mean(u1)
  fhat=mean(d1)-mean(d2)
  phat=mean(o1)-mean(o2)
  sscarry=2*n1*n2/(n1+n2)*chat^2/4
  ssdrug=2*n1*n2/(n1+n2)*fhat^2
  ssperiod=2*n1*n2/(n1+n2)*phat^2
  ssintra=sswit-ssdrug-ssperiod
  ssinter=ssbet-sscarry
  
  source1=c("carry","inter","drug","period","intra","Total") #source of var
  dof=c(1,n1+n2-2,1,1,n1+n2-2,n-1)
  ss=c(sscarry,ssinter,ssdrug,ssperiod,ssintra,sst)
  msq=ss/dof
  fcal=rep(0,6)
  fcal[1]=msq[1]/msq[2]
  fcal[2:4]=msq[2:4]/msq[5]
  pval=rep(0,6)
  pval[1]=1-pf(fcal[1],dof[1],dof[2])
  pval[2:4]=1-pf(fcal[2:4],dof[2:4],dof[5])
  return(data.frame("s_var"=source1,"DF"=dof,"sof_squ"=round(ss,2),"Fcal"=round(fcal,2),"p_value"=round(pval,4)))
}
```

```{r,echo=FALSE}
#datasets:
data_1=read.csv("E:/PRACTICAL SEM III/data set/BABEData22.csv")
df=data.frame(data_1,out)
df
```

```{r,echo=FALSE}
# Anova corresponds to concentration:
crosanova(df$cmax,df$Seq,df$Sub,df$Per)
```

# Interpretation:

-   carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}\neq0$, here from p_value we fail to reject $H_01$ at 5% level of significance (i.e. Carryover effect of Test drug and Reference drug is same).
-   from p_value we can conclude that inter subject variability is present.
-   Drug effect $H_{03}:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}\neq0$ here from p_value we fail to reject $H_{03}$ at 5% level of significance (i.e. Drug effect of Test drug and Reference drug is same).
-   period effect $H_{04}:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}\neq0$, here from p_value fail to reject $H_{04}$ at 5% level of significance (i.e. Period effect of Test drug and Reference drug is same).
-   from p_value we can conclude that intra subject variability is not present.

```{r}
# Anova corresponds to time point:
crosanova(df$tmax,df$Seq,df$Sub,df$Per)
```

# Interpretation:

-   carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ not equal to 0 here from p_value we fail to reject $H_01$ at 5% level of significance (i.e. Carryover effect of Test drug and Reference drug is same).
-   from p_value we can conclude that inter subject variability is present.
-   Drug effect $H_{03}:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}\neq0$, here from p_value we fail to reject $H_03$ at 5% level of significance (i.e. Drug effect of Test drug and Reference drug is same).
-   period effect $H_{04}:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}\neq0$, here from p_value fail to reject $H_{04}$ at 5% level of significance.(i.e. Period effect of Test drug and Reference drug is same).
-   from p_value we can conclude that intra subject variability is not present.

```{r}
# Anova corresponds to area under curve 
crosanova(df$auct,df$Seq,df$Sub,df$Per)
```

\#**Interpretation :**

-   carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}\neq0$ here from p_value we fail to reject $H_01$ at 5% level of significance (i.e. Carryover effect of Test drug and Reference drug is same).
-   from p_value we can conclude that inter subject variability is present.
-   Drug effect $H_{03}:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}\neq0$, here from p_value we fail to reject $H_03$ at 5% level of significance (i.e. Drug effect of Test drug and Reference drug is same).
-   period effect $H_{04}:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}\neq0$, here from p_value fail to reject $H_{04}$ at 5% level of significance (i.e. Period effect of Test drug and Reference drug is same).
-   from p_value we can conclude that intra subject variability is not present.

```{r}
# Anova corresponds to time point concentration max
crosanova(df$ke,df$Seq,df$Sub,df$Per)
```

\#**Interpretation :**

-   carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}\neq0$, here from p_value we fail to reject $H_01$ at 5% level of significance (i.e. Carryover effect of Test drug and Reference drug is same).
-   from p_value we can conclude that inter subject variability is present.
-   Drug effect $H_{03}:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}\neq0$, here from p_value we fail to reject $H_{03}$ at 5% level of significance (i.e. Drug effect of Test drug and Reference drug is same).
-   Period effect $H_{04}:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}\neq0$, here from p_value we fail to reject $H_{04}$ at 5% level of significance (i.e. Period effect of Test drug and Reference drug is same).
-   from p_value we can conclude that intra subject variability is not present.

```{r}
# Anova corresponds to partial auc truncated at the population median of t max 
crosanova(df$aucinf,df$Seq,df$Sub,df$Per)
```

\#**Interpretation :**

-   Carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}$ here from p_value we fail to reject $H_01$ at 5% level of significance (i.e. Carryover effect of Test drug and Reference drug is same).
-   from p_value we can conclude that inter subject variability is present.
-   Drug effect $H_{03}:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}\neq0$, here from p_value we fail to reject $H_03$ at 5% level of significance (i.e. Drug effect of Test drug and Reference drug is same).
-   period effect $H_{04}:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}\neq0$, here from p_value we fail to reject $H_{04}$ at 5% level of significance (i.e. Period effect of Test drug and Reference drug is same).
-   from p_value we can conclude that intra subject variability is not present.

```{r}
# Anova corresponds to elimination half time 
crosanova(df$thalf,df$Seq,df$Sub,df$Per)
```

#Interpretation:

\- carryover effect $H_{01}:{C_T}-{C_R}=0$ vs $H_{11}:{C_T}-{C_R}\neq0$ here from p_value we fail to reject $H_01$ at 5% level of significance (i.e. Carryover effect of Test drug and Reference drug is same).

-   from p_value we can conclude that inter subject variability is present.
-   Drug effect $H_{03}:{F_T}-{F_R}=0$ vs $H_{13}:{F_T}-{F_R}\neq0$, here from p_value we fail to reject $H_03$ at 5% level of significance (i.e. Drug effect of Test drug and Reference drug is same).
-   period effect $H_{04}:{P_1}-{P_2}=0$ vs $H_{14}:{P_1}-{P_2}\neq0$, here from p_value we fail to reject $H_{04}$ at 5% level of significance (i.e. Period effect of Test drug and Reference drug is same).
-   from p_value we can conclude that intra subject variability is not present.

```{r,echo=FALSE}
#merging of two graph:
time=unlist(ti)
y1=apply(data[data$V3==1,5:20],2,mean)
x1=plot(time,y1,type="l",col='blue',main="Area Under Curve",xlab = "time",ylab="concentration")
par(new=TRUE)
y2=apply(data[data$V3==2,5:20],2,mean)
x2=plot(time,y2,type="l",col='red',main ="Area Under Curve",xlab = "time",ylab="concentration")
```
